Making Deep Q-learning Methods Robust to Time Discretization

### 2019-01-19 paper/icml_drau/advup.tex
experiments
relecture advup

\subsection{Advantage Updating}
\label{subsec:reparam}
To define an object which contains the same information on actions as
$Q^\pi_\deltat$, but admits an action dependant continuous time limit, it is
natural to define \cite{adv_upd}
\begin{align}
	A^\pi_\deltat(s, a) &\mathdef \frac{Q^\pi_\deltat(s,a) - V^\pi_\deltat(s)}{\deltat},
    \label{eq:adv}
\end{align}
a rescaled version of the advantage function \cite{adv_upd}, as the difference between between
$Q^\pi_\deltat(s, a)$ and $V^\pi\deltat(s)$ is of order $\bigO(\deltat)$.
Contrary to the action state value function, the rescaled advantage function converges when $\deltat$ goes to $0$
to an action dependent quantity which contains the necessary information for policy improvement:

@@

%\TODO{CT and LB are fighting each other about the importance which should be given to the next paragraph. A impartial judge should give his opinion. (CT is wrong)}
However, the proposed reparameterization does not, on its own, guarantee that
when $Q$ correctly approximates $Q^\pi_\deltat$, $A$ approximates $A^\pi_\deltat$\footnote{
	Ensuring that $A$ do not differ from $A^\pi_\deltat$ by a state dependent function only affects
	$A$'s interpretability. Indeed, for any function $f$, the ranking of actions given by
	$A^\pi_\deltat(s, \cdot)$ and $A^\pi_\deltat(s, \cdot) + f(s)$ is the same.
}.
Indeed, for any given pair $V$, $A$,
\begin{align}
	\tilde{V}(s) &= V(s) - f(s)\\
	\tilde{A}(s, a) &= A(s, a) + \frac{f(s)}{\deltat}
\end{align}
yields the exact same $Q$ function. Consequently, $Q$ can correctly approximate
$Q^\pi$ even if $A$ is arbitrarily far from $A^\pi$.
To enforce identifiability of $A$, one must enforce the consistency equation
\begin{equation}
	Q^\pi_\deltat(s, \pi(s)) - V^\pi_\deltat(s) = 0
\end{equation}
on the approximate $A$ and $V$. This translates to
\begin{equation}
	A(s, \pi(s)) = 0.
\end{equation}
With this additional constraint, if $Q = Q^\pi_\deltat$, then $A = A^\pi_\deltat$ and $V = V^\pi_\deltat$.
\begin{align}
	A^\pi_\deltat(s, a) &= \frac{Q^\pi_\deltat(s,a) - V^\pi_\deltat(s)}{\deltat}\\
		    &= \frac{Q(s,a) - Q(s, \pi(s))}{\deltat}\\
		    &= A(s, a).
\end{align}
In the spirit of~\cite{dueling_nets}, this condition is enforced by writing $A$ as
\begin{equation}
\label{eq:Aparam}
	A(s, a) = \bar{A}(s, a) - \bar{A}(s, \pi(s)).
\end{equation}
With this parameterization, one can hope to learn an approximation of $A^\pi_\deltat$,
from which a proper hiearchy of actions can be derived for any $\deltat$.

\subsection{Time invariant exploration}
\label{subsec:explo}
To obtain a time invariant reinforcement learning algorithm, a time invariant
exploration scheme is required. 
\cite{ddpg} already introduces an exploration scheme for continuous actions, which consists in adding an
\emph{Ornstein-Uhlenbeck}~\cite{orn-uhl} process (OU process) to actions, that is naturally time
discretization invariant. Formally, it is defined as
\begin{equation}
	\pi^\text{explore}(s^k_\deltat, z^k_\deltat) = \pi(s^k_\deltat) + z^k_\deltat
\end{equation}
with $z^k_\deltat$ the discretization of a time continuous OU process, defined as
\begin{equation}
	dz_t = - z_t \theta dt + \sigma \dbt.
	\label{eq:orn_uhl}
\end{equation}
where $B_t$ is a brownian motion. Resulting discretized trajectories converge
to non trivial continuous time trajectories.

@@

Contrary to this discrete action exploration scheme, $\varepsilon$-greedy
exploration is likely to not explore, i.e. to collapse to a deterministic
policy when $\deltat$ goes to $0$. Indeed, one can prove that in a near
continuous deterministic environment, the exploration policy resulting from
applying $\varepsilon$ greedy exploration to a constant exploitation policy
converges to a determinist policy as $\deltat$ goes to $0$. More precisely, let
$(a_\deltat)$ a sequence of i.i.d. actions of distribution $p_A$, (for instance
the sequence of actions generated by $\varepsilon$-greedy exploration on a
constant policy). Then, as $\deltat$ goes to $0$, trajectories generated by
interacting with the environment using the $a$'s converge to solutions of
\begin{equation}
	ds_t = \E_{a\sim p_A}\left[F(s_t, a)\right]dt
\end{equation}
where the expectation cancels the exploration noise. The proof is given in the
supplementary material.

\subsection{Algorithms}
\label{subsec:algorithm}
\begin{algorithm}[ht]
	\caption{Deep Advantage Updating (Discrete actions)}
	\input{dauc.tex}
	\label{alg:dau}
\end{algorithm}

@@

For parameter updates to follow a SDE, one must
have $\eta^V_\deltat = \alpha^V \deltat$ and $\eta^A_\deltat = \alpha^A
\deltat$. More precisely
\begin{theorem}
	\TODO{We need a bit more details: we probably want to assume that states and actions are
	drawn from a single fixed trajectory, and that we are learning for a fixed, constant policy.}
	Let $\eta^V_\deltat = \alpha^V \deltat^\beta$ (resp. $\eta^A \deltat^\beta$) be the learning rate
	of $V$ (resp. $A$), when training on a single exploratory trajectory. Then
	\begin{itemize}
		\item If $\beta = 1$ the discrete parameter trajectories converge to continuous parameter
			trajectories as $\deltat$ goes to $0$.
		\item If $\beta < 1$, parameters grow arbitrarily large in a single timestep as $\deltat$
			goes to $0$.
		\item If $\beta > 1$, parameter trajectories shrink to a single point as 
			$\deltat$ goes to $0$.
	\end{itemize}
	\label{th:cont-params}
\end{theorem}

>>>>

\subsection{Advantage Updating}
\label{subsec:reparam}
To define an object which contains the same information on actions as
$Q^\pi_\deltat$, but admits an action dependant continuous time limit, it is
natural to define \cite{adv_upd}
\begin{align}
	A^\pi_\deltat(s, a) &\mathdef \frac{Q^\pi_\deltat(s,a) - V^\pi_\deltat(s)}{\deltat},
    \label{eq:adv}
\end{align}
a rescaled version of the advantage function, as the difference between between
$Q^\pi_\deltat(s, a)$ and $V^\pi_\deltat(s)$ is of order $\bigO(\deltat)$.
Contrary to the action state value function, the rescaled advantage function converges when $\deltat$ goes to $0$
to an action dependent quantity which contains the necessary information for policy improvement:

@@

%\TODO{CT and LB are fighting each other about the importance which should be given to the next paragraph. A impartial judge should give his opinion. (CT is wrong)}
However, the proposed reparameterization does not, on its own, guarantee that
when $Q$ correctly approximates $Q^\pi_\deltat$, $A$ approximates $A^\pi_\deltat$\footnote{
	Ensuring that $A$ does not differ from $A^\pi_\deltat$ by a state dependent function only affects
	$A$'s interpretability. Indeed, for any function $f$, the ranking of actions given by
	$A^\pi_\deltat(s, \cdot)$ and $A^\pi_\deltat(s, \cdot) + f(s)$ is the same.
}.
Indeed, for any given pair $V$, $A$,
\begin{align}
	\tilde{V}(s) &= V(s) - f(s)\\
	\tilde{A}(s, a) &= A(s, a) + \frac{f(s)}{\deltat}
\end{align}
yields the exact same $Q$ function. Consequently, $Q$ can correctly approximate
$Q^\pi$ even if $A$ is arbitrarily far from $A^\pi$.
To enforce identifiability of $A$, one must enforce the consistency equation
\begin{equation}
	Q^\pi_\deltat(s, \pi(s)) - V^\pi_\deltat(s) = 0
\end{equation}
on the approximate $A$ and $V$. This translates to
\begin{equation}
	A(s, \pi(s)) = 0.
\end{equation}
With this additional constraint, if $Q = Q^\pi_\deltat$, then $A = A^\pi_\deltat$ and $V = V^\pi_\deltat$.
\begin{align}
	A^\pi_\deltat(s, a) &= \frac{Q^\pi_\deltat(s,a) - V^\pi_\deltat(s)}{\deltat}\\
		    &= \frac{Q(s,a) - Q(s, \pi(s))}{\deltat}\\
		    &= A(s, a).
\end{align}
In the spirit of~\cite{dueling_nets}, this condition is enforced by writing $A$ as
\begin{equation}
\label{eq:Aparam}
	A(s, a) = \bar{A}(s, a) - \bar{A}(s, \pi(s)).
\end{equation}
With this parameterization, one can hope to learn an approximation of $A^\pi_\deltat$,
from which a proper ranking of actions can be derived for any $\deltat$.

\subsection{Time invariant exploration}
\label{subsec:explo}
To obtain a time invariant RL algorithm, a time invariant
exploration scheme is required. 
\cite{ddpg} already introduces a time discretization invariant exploration scheme for continuous actions, which consists in adding an
\emph{Ornstein-Uhlenbeck}~\cite{orn-uhl} process (OU process) to actions. Formally, it is defined as
\begin{equation}
	\pi^\text{explore}(s^k_\deltat, z^k_\deltat) = \pi(s^k_\deltat) + z^k_\deltat
\end{equation}
with $z^k_\deltat$ the discretization of a time continuous OU process, 
\begin{equation}
	dz_t = - z_t \kappa dt + \sigma \dbt.
	\label{eq:orn_uhl}
\end{equation}
where $B_t$ is a brownian motion, $\kappa$ a stiffness parameter and $\sigma$ a noise scaling parameter. Resulting discretized trajectories converge
to non trivial continuous time trajectories.

@@

Contrary to this discrete action exploration scheme, $\varepsilon$-greedy
exploration is likely to not explore, i.e. to collapse to a deterministic
policy when $\deltat$ goes to $0$. Indeed, one can prove that in a near
continuous deterministic environment, the exploration policy resulting from
applying $\varepsilon$-greedy exploration to a constant exploitation policy
converges to a determinist policy as $\deltat$ goes to $0$. More precisely, let
$(a_\deltat)$ be a sequence of i.i.d. actions of distribution drawn from $p_A$, (for instance
the sequence of actions generated by $\varepsilon$-greedy exploration on a
constant policy). Then, as $\deltat$ goes to $0$, trajectories generated by
interacting with the environment using the $a$'s converge to solutions of
\begin{equation}
	ds_t = \E_{a\sim p_A}\left[F(s_t, a)\right]dt
\end{equation}
where the expectation cancels the exploration noise. The proof is given in the
supplementary material. \TODO{Is there a stronger theorm? Should we talk more about that?}

\subsection{Algorithms}
\label{subsec:algorithm}
\begin{algorithm}[ht]
  \caption{Deep Advantage Updating (Discrete actions)}
  \label{alg:dau}
	\input{dauc.tex}
\end{algorithm}

@@

For parameter updates to follow a SDE, one must
have $\eta^V_\deltat = \alpha^V \deltat$ and $\eta^A_\deltat = \alpha^A
\deltat$. More precisely
\begin{theorem}
	\TODO{We need a bit more details: we probably want to assume that states and actions are
	drawn from a single fixed trajectory, and that we are learning for a fixed, constant policy.}
	Let $\eta^V_\deltat = \alpha^V \deltat^\beta$ (resp. $\eta^A \deltat^\beta$) be the learning rate
	of $V$ (resp. $A$), when training on a single exploratory trajectory. Then
	\begin{itemize}
		\item If $\beta = 1$ the discrete parameter trajectories converge to continuous parameter
			trajectories as $\deltat$ goes to $0$.
		\item If $\beta < 1$, parameters grow arbitrarily large in a single timestep as $\deltat$
			goes to $0$.
		\item If $\beta > 1$, parameter trajectories shrink to a single point as 
			$\deltat$ goes to $0$. \TODO{Is that clear}
	\end{itemize}
	\label{th:cont-params}
\end{theorem}
